---
title: "Tidy Tuesday Week 20"
author: "Alex Kinkelaar"
date: "5/16/2023"
output: html_document
---

```{r opening, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidy Tuesday Week 20

### Setup: Loading Data and Necessary Packages

```{r Setup}
library(tidyverse)
library(tidytuesdayR)
library(usmap)
library(alex)
library(png)
data = tidytuesdayR::tt_load(2023, week = 20)
data = data$tornados
```

### A Basic Solution

Because I'm chosing to use the `usmap` package, I'll need to get the data I want to map into a format that package can plot. `usmap` looks for a `fips` column to plot counties against - I'll start by plotting just the first county each tornado is reported in, so a simple rename is all thats necessary. The 5 digit FIPS code doesn't exist in this dataset, so I'll need to assemble it from the county codes and the state codes. I'll then want to summarize that data to get the counts of tornados that started in each county.

```{r Basic Plot}
plot_data = data %>% 
  filter(f1 > 0) %>%
  mutate(t_id = paste(yr, om, sep = "-")) %>%
  mutate(three_f1 = str_pad(string = f1, width = 3, pad = "0")) %>%
  select(three_f1, stf, t_id) %>%
  mutate(two_stf = str_pad(string = stf, width = 2, pad = "0")) %>%
  mutate(fips = paste(two_stf, three_f1, sep = "")) %>%
  group_by(fips) %>%
  summarize(tornados = n())

plot_map = plot_usmap(
  regions = "counties",
  data = plot_data,
  values = "tornados",
  size = 0.1) +
  labs(title = "Number of Tornados", 
       subtitle = "By U.S. County, 1950-2022", 
       caption = "Tornados only counted in first county reported") +
  scale_fill_continuous(name = "Tornados") +
  theme_void() +
  theme(
    legend.position = "right",
    plot.background = element_rect(fill = "#262626", color = NA),
    legend.text = element_text(color = "white"),
    legend.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.caption = element_text(face = "italic", size = 10, color = "white"),
    plot.title = element_text(face = "bold", size = 14, color = "white"),
    plot.subtitle = element_text(face = "italic", size = 11, color = "white"),
    plot.margin = unit(c(1,1,1,1), "cm")
  )

plot_map

ggsave(filename = "plot_map.png",
       plot = plot_map)
```





